---
title: "Implementation"
author: "Simon Adamov"
date: "August 12, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      error = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.retina =3,
                      fig.width = 10,
                      fig.height = 7,
                      out.width = "100%",
                      out.height = "100%")

library(mchdwh)
library(dplyr)
library(lubridate)
library(purrr)

# This library is needed for visual representation of missing data
# library(Amelia)
# remotes::install_version("foreign", version = "0.8-71")
# install.packages("Amelia")

# These themes just look great :-)
library(ggthemr)
# devtools::install_github('cttobin/ggthemr')
ggthemr("fresh")
```

# Data Import

The data is obtained from the DWH for the current pollen season (2020). 

```{r }

# The following types are being measured:
# Erle - Alder - Aulne - Alnus
# Birke - Birch - Bouleau - Betula
# Gräser - Grasses - Graminées - Poaceae
# Ambrosia - Ragweed - Ambroisie - Ambrosia

species_all <- tibble(
  taxon = c(
    "Castanea",
    "Alnus",
    "Ulmus",
    "Cupressus",
    "Fraxinus",
    "Fagus",
    "Juglans",
    "Plantago",
    "Corylus",
    "Pinus",
    "Quercus",
    "Rumex",
    "Platanus",
    "Populus",
    "Poaceae",
    "Salix",
    "Betula",
    "Carpinus",
    "Urtica",
    "Taxus",
    "Picea",
    "Ambrosia"
  ),
  tag = c(
    "kacastz0",
    "kaalnuz0",
    "kaulmuz0",
    "kacuprz0",
    "kafraxz0",
    "kafaguz0",
    "kajuglz0",
    "khplanz0",
    "kacoryz0",
    "kapinuz0",
    "kaquerz0",
    "khrumez0",
    "kaplatz0",
    "kapopuz0",
    "khpoacz0",
    "kasaliz0",
    "kabetuz0",
    "kacarpz0",
    "khurtiz0",
    "kataxuz0",
    "kapicez0",
    "khambrz0"
  )
)

species <- species_all %>%
  filter(taxon %in% c("Alnus", "Ambrosia", "Betula", "Poaceae"))

stations <-
  tibble(
    tag = c(
      "PDS",
      "PBU",
      "PMU",
      "PBS",
      "PZH",
      "PLZ",
      "PBE",
      "PPY",
      "PNE",
      "PVI",
      "PLS",
      "PGE",
      "PCF",
      "PLO",
      "BLR",
      "PLU"
    ),
    commune = c(
      "Wolfgang",
      "Buchs",
      "Münsterlingen",
      "Basel",
      "Zürich",
      "Luzern",
      "Bern",
      "Payerne",
      "Neuchâtel",
      "Visp",
      "Lausanne",
      "Genève",
      "La-Chaux-de-Fonds",
      "Locarno",
      "Balerna",
      "Lugano"
      )
  )


```

```{r eval = FALSE}
# Loading the Data takes a long time and often fails when multiple years are retrieved.
# data_hirst <- dwhget_surface(param_short = species$tag,
#                                   meas_cat = 204,
#                                   year = 2016:2020,
#                                   iso_country_cd = "CH") %>% 
#   mutate(datetime = ymd_hm(datetime)) %>% 
#   mutate_if(is.character, ~as.factor(.)) %>% 
#   inner_join(species, by = c("param_short" = "tag")) %>% 
#   inner_join(stations, by = c("nat_abbr" = "tag"))
# 
# save(data_hirst, file = paste0(here::here(), "/vignettes/ext-data/data_hirst.RData"))

load(paste0(here::here(), "/vignettes/ext-data/data_hirst.RData"))
```


# Data Bias Correction

The team in Payerne found that the air sucking rates of the Hirst traps are actually higher in reality than reported by the manufacturer. The traps suck in 13.5 l per minute instead of 10 l per minute. Hence our Pollen counts per 10 minutes are too high and should be reduced by a factor of 1.35.

```{r eval = FALSE}
data_hirst<- data_hirst %>% 
  mutate(value_unbiased = value / 1.35)
```

# Missing Data and Imputation

There are no missing measurements.
Generally, missing data usually occurs when the silicone rubber bands were being exchanged.
Are they being imputed before storing the data on the DWH?

```{r  eval = FALSE}

# This takes ~ 2minutes - visual representation of missing data
# data_hirst %>% 
#   as.data.frame() %>% 
#   missmap(y.labels = "", y.at = 1, col = swatch()[c(4,2)])

data_hirst %>% 
  summarise_all(~sum(is.na(.)))

summary(data_hirst)

```


# Temporal Aggregation

There are some general guidelines from Regula Gehrig on how to average concentration that should be followed here.

Die Stundenwerte werden wie folgt aus den 10-Minutenwerten berechnet:
z.B. Wert für 6:00 Uhr UTC wird berechnet aus 05:10 – 6:00 UTC

Die 3-h-Werte, die du auch im DWH findest, werden so berechnet:
z.B. Wert um 09:00 UTC: 06:10-09:00

Die Tageswerte werden bei den Pollen glaub ich aus den Stundenwerten berechnet aus:
01:00 – 24:00 (d.h. 00:00 – im DWH gibt es keinen Wert 24:00) UTC 


```{r eval = FALSE}

data_hirst <- data_hirst %>%
  mutate(date = date(datetime),
         hms = format(datetime, "%H:%M:%S"),
         hour = lag(hour(datetime), 1) + 1,
         hour = ifelse(hour == 24, 0, hour))
  
data_hirst_window <- data_hirst %>% 
  filter(between(datetime, as_datetime("2017-03-02 00:10:00"), as_datetime("2020-08-19 00:00:00"))) # Only look at full days

data_hourly <- data_hirst_window
         
data_2hours <- data_hirst_window %>%
      mutate(hour = case_when(
      hms > "00:00:00" & hms <= "02:00:00" ~ 2,
      hms > "02:00:00" & hms <= "04:00:00" ~ 4,
      hms > "04:00:00" & hms <= "06:00:00" ~ 6,
      hms > "06:00:00" & hms <= "08:00:00" ~ 8,
      hms > "08:00:00" & hms <= "10:00:00" ~ 10,
      hms > "10:00:00" & hms <= "12:00:00" ~ 12,
      hms > "12:00:00" & hms <= "14:00:00" ~ 14,
      hms > "14:00:00" & hms <= "16:00:00" ~ 16,
      hms > "16:00:00" & hms <= "18:00:00" ~ 18,
      hms > "18:00:00" & hms <= "20:00:00" ~ 20,
      hms > "20:00:00" & hms <= "22:00:00" ~ 22,
      hms > "22:00:00" | hms == "00:00:00" ~ 0 # For the next day already
    ))

data_3hours <- data_hirst_window %>%
  mutate(
    hour = case_when(
      hms > "00:00:00" & hms <= "03:00:00" ~ 3,
      hms > "03:00:00" & hms <= "06:00:00" ~ 6,
      hms > "06:00:00" & hms <= "09:00:00" ~ 9,
      hms > "09:00:00" & hms <= "12:00:00" ~ 12,
      hms > "12:00:00" & hms <= "15:00:00" ~ 15,
      hms > "15:00:00" & hms <= "18:00:00" ~ 18,
      hms > "18:00:00" & hms <= "21:00:00" ~ 21,
      hms > "21:00:00" | hms == "00:00:00" ~ 0 # For the next day already
    ))

data_6hours <- data_hirst_window %>%
  mutate(
    hour = case_when(
      hms > "00:00:00" & hms <= "06:00:00" ~ 6,
      hms > "06:00:00" & hms <= "12:00:00" ~ 12,
      hms > "12:00:00" & hms <= "18:00:00" ~ 18,
      hms > "18:00:00" | hms == "00:00:00" ~ 0 # For the next day already
    ))

data_12hours <- data_hirst_window %>%
  mutate(
    hour = case_when(
      hms > "00:00:00" & hms <= "12:00:00" ~ 12,
      hms > "12:00:00" | hms == "00:00:00" ~ 0 # For the next day already
    ))

data <- map(list(data_hourly, data_2hours, data_3hours, data_6hours, data_12hours), ~.x %>% 
  mutate(date = if_else(hour == 0 & hms != "00:00:00", date + days(1), date)) %>% # This way they will be allocated to the next day
  group_by(taxon, commune, date, hour) %>% 
  summarise(value = mean(value_unbiased)) %>% 
  ungroup() %>% 
  mutate(datetime = ymd_hm(paste0(as.character(date), paste0(sprintf("%02d", hour), ":00")))))

data[[6]] <- data[[1]] %>% 
  group_by(date, commune, taxon) %>% 
  summarise(value = mean(value)) %>% 
  ungroup() %>% 
  mutate(datetime = date,
         hour = 00:00)

names(data) <- c("data_hourly", "data_2hours", "data_3hours", "data_6hours", "data_12hours", "data_daily")

# save(data, file = paste0(here::here(), "/vignettes/ext-data/data_aggs.RData"))
  
``` 

```{r}
load(paste0(here::here(), "/vignettes/ext-data/data_aggs.RData"))
```


# Time Series Plots

```{r fig.height=9, fig.width=16}
stations_subset <- map(c(4, 8, 12, 16), ~stations %>% 
                         arrange(commune) %>% 
                         slice((.x - 3):.x) %>% 
                         pull(commune))

map(stations_subset, ~data$data_daily %>% 
  filter(commune %in% .x) %>% 
  ggplot(aes(x = datetime, y = value, col = taxon)) +
    geom_line() +
    # coord_cartesian(ylim = c(0, 5000)) +
    labs(y = expression(paste("Pollen Concentration [", m^-3, "]")),
       x = "") +
  scale_colour_discrete("") +
    facet_wrap(~commune))
```


```{r}

commune <- "Payerne" # Check stations tibble for all available locations
taxon <- "Betula" # Check species tibble for all available pollen taxa

data$data_daily %>% 
  filter(taxon == taxon,
         commune == commune) %>% 
  ggplot(aes(x = datetime, y = value, col = taxon)) +
    geom_line(alpha = 0.6) +
    ggtitle(paste("Hourly Pollen Concentrations in", commune)) +
  labs(y = expression(paste("Pollen Concentration [", m^-3, "]")),
       x = "") +
  scale_colour_discrete("")
```














































